---
title: "Data Visualization"
author: "William Donovan"
output: html_document
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggridges)
library(p8105.datasets)
```

Import weather data

```{r}
data("weather_df")
```

Make a simple scatter plot

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()
```

Now add in color for the location

```{r}
# Can also pipe in the dataframe for a different style
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point()
```

Change the opacity of the points

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) # Set the alpha for opacity
```

Add a curve/line

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  geom_smooth(se = FALSE) # add geom_smooth for a nice line
```

Change where the color is assigned so that only one line appears

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(alpha = 0.5, aes(color = name)) + # Color added ONLY to geom_point
  geom_smooth(se = FALSE)
```

Show preciptitation based on SIZE

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(alpha = 0.5, aes(color = name, size = prcp)) + # add size to change the point size
  geom_smooth(se = FALSE)
```

Now facet the plots in order to better visualize

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.4, aes(size = prcp)) +
  geom_smooth(se = FALSE) +
  facet_grid((name ~ .)) # ". ~ name" would produce the plots in a row rather than a column
```

### Learning Assesment

```{r}
weather_df |>
  filter(name == "Waterhole_WA") |>
  mutate(
    tmin_f = (5/9)*(tmin-32),
    tmax_f = (5/9)*(tmax-32)) |>
  ggplot(aes(x = tmin_f, y = tmax_f)) +
  geom_point(alpha = 0.4, aes(size = prcp)) +
  geom_smooth(se = FALSE, method = "lm", color = "black")
```

Geom_hex is kinda cool

```{r}
weather_df |>
  filter(name == "Waterhole_WA") |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()
```

Density plot

```{r}
weather_df |>
  ggplot(aes(x = tmax)) +
  geom_density(aes(fill = name), alpha = 0.4)
```

Violin plot

```{r}
weather_df |>
  ggplot(aes(x = name, y = tmax)) +
  geom_violin(aes(fill = name), alpha = 0.6)
```

Ridge plot

```{r}
weather_df |>
  ggplot(aes(x = tmax, y = name)) +
  geom_density_ridges(aes(fill = name), alpha = 0.8)
```

### Learning Assesment maybe?

```{r}
weather_df |>
  filter(date <= "2021-12-31") |>
  filter(prcp != 0) |>
  ggplot(aes(x = date, y = prcp)) +
  geom_point(aes(color = name), alpha = 0.5) +
  facet_grid(name ~ .)
```

# Data Visualization Lecture 2

## More advanced stuff

### Add labels with labs()

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name)) +
  labs(
    x = "Minimum daily temperature (C)",
    y = "Maximum daily temperature (C)",
    title = "Temperature Ranges",
    color = "Location",
    caption = "Data from NOAA"
  )
```

### Adjust the scales

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name)) +
  labs(
    x = "Minimum daily temperature",
    y = "Maximum daily temperature (C)",
    title = "Temperature Ranges",
    color = "Location",
    caption = "Data from NOAA") +
  scale_x_continuous(
    breaks = c(-15, 0, 15),
    labels = c("-15 C", "0 C", "15 C")
  )
```

### Change color using Viridis package

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name)) +
  labs(
    x = "Minimum daily temperature",
    y = "Maximum daily temperature (C)",
    title = "Temperature Ranges",
    color = "Location",
    caption = "Data from NOAA") +
  scale_x_continuous(
    breaks = c(-15, 0, 15),
    labels = c("-15 C", "0 C", "15 C")) +
  viridis::scale_color_viridis(
    discrete = TRUE,
    alpha = 0.4)
```

### Move the legend

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name)) +
  labs(
    x = "Minimum daily temperature",
    y = "Maximum daily temperature (C)",
    title = "Temperature Ranges",
    color = "Location",
    caption = "Data from NOAA") +
  scale_x_continuous(
    breaks = c(-15, 0, 15),
    labels = c("-15 C", "0 C", "15 C")) +
  viridis::scale_color_viridis(
    discrete = TRUE,
    alpha = 0.7) +
  theme(legend.position = "bottom")
```

### Save plot as an object and change the theme

```{r}
ggp_temp_scatter = 
  weather_df |>
    ggplot(aes(x = tmin, y = tmax)) +
    geom_point(aes(color = name)) +
    labs(
      x = "Minimum daily temperature",
      y = "Maximum daily temperature (C)",
      title = "Temperature Ranges",
      color = "Location",
      caption = "Data from NOAA") +
    scale_x_continuous(
      breaks = c(-15, 0, 15),
      labels = c("-15 C", "0 C", "15 C")) +
    viridis::scale_color_viridis(
      discrete = TRUE,
      alpha = 0.7)
    

ggp_temp_scatter +
  theme_bw() +
  theme(legend.position = "bottom")
```

## Patchwork

### Make three plots and combine using patchwork

```{r}
library(patchwork)

ggp_prcp_density = 
  weather_df |>
    filter(prcp > 0) |>
    ggplot(aes(x = prcp)) +
    geom_density(aes(fill = name), alpha = 0.5)

ggp_prcp_density

(weather_scatterplot + ggp_prcp_density)

```


## New Challenge

```{r}
pups_df = read_csv("data/FAS_pups.csv",
                   na = c("NA", "", " ", "."),
                   skip = 3) |>
  janitor::clean_names() |>
  mutate(sex = case_match(
    sex,
    1 ~ "Male",
    2 ~ "Female"))

pulse_df = haven::read_sas("data/public_pulse_data.sas7bdat")

litters_df = read_csv("data/FAS_litters.csv",
                   na = c("NA", "", " ", ".")) |>
  janitor::clean_names() |>
  separate(group, into = c("dose", "tx_day"), sep = 3)

fas_df = left_join(pups_df, litters_df, by = "litter_number")

fas_df = fas_df |>
  select(pd_ears:tx_day) |>
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome",
    names_prefix = "pd_",
    values_to = "pn_day") |>
  drop_na()

View(fas_df)

```

