---
title: "Data Visualization"
author: "William Donovan"
output: html_document
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggridges)
library(p8105.datasets)
```

Import weather data

```{r}
data("weather_df")
```

Make a simple scatter plot

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()
```

Now add in color for the location

```{r}
# Can also pipe in the dataframe for a different style
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point()
```

Change the opacity of the points

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) # Set the alpha for opacity
```

Add a curve/line

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  geom_smooth(se = FALSE) # add geom_smooth for a nice line
```

Change where the color is assigned so that only one line appears

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(alpha = 0.5, aes(color = name)) + # Color added ONLY to geom_point
  geom_smooth(se = FALSE)
```

Show preciptitation based on SIZE

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(alpha = 0.5, aes(color = name, size = prcp)) + # add size to change the point size
  geom_smooth(se = FALSE)
```

Now facet the plots in order to better visualize

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.4, aes(size = prcp)) +
  geom_smooth(se = FALSE) +
  facet_grid((name ~ .)) # ". ~ name" would produce the plots in a row rather than a column
```

### Learning Assesment

```{r}
weather_df |>
  filter(name == "Waterhole_WA") |>
  mutate(
    tmin_f = (5/9)*(tmin-32),
    tmax_f = (5/9)*(tmax-32)) |>
  ggplot(aes(x = tmin_f, y = tmax_f)) +
  geom_point(alpha = 0.4, aes(size = prcp)) +
  geom_smooth(se = FALSE, method = "lm", color = "black")
```

Geom_hex is kinda cool

```{r}
weather_df |>
  filter(name == "Waterhole_WA") |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()
```

Density plot

```{r}
weather_df |>
  ggplot(aes(x = tmax)) +
  geom_density(aes(fill = name), alpha = 0.4)
```

Violin plot

```{r}
weather_df |>
  ggplot(aes(x = name, y = tmax)) +
  geom_violin(aes(fill = name), alpha = 0.6)
```

Ridge plot

```{r}
weather_df |>
  ggplot(aes(x = tmax, y = name)) +
  geom_density_ridges(aes(fill = name), alpha = 0.8)
```

### Learning Assesment maybe?

```{r}
weather_df |>
  filter(date <= "2021-12-31") |>
  filter(prcp != 0) |>
  ggplot(aes(x = date, y = prcp)) +
  geom_point(aes(color = name), alpha = 0.5) +
  facet_grid(name ~ .)
```




